### 2022-03-19 outputs/paper/paper.Rmd
first commit test

---
title: "My titleXXXXXXXXXXXXXXXX"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(palmerpenguins)
```

# Introduction

You can and should cross-reference sections and sub-sections. For instance, Section \@ref(data). R Markdown automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point).

# Data

Our data is of penguins (Figure \@ref(fig:bills)).

```{r bills, fig.cap="Bills of penguins", echo = FALSE}
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)

```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=8, fig.height=4}

# This needs to be about the random data tha tI sasmpled?
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk way more about it.

# Model

\begin{equation}
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}  (\#eq:bayes)
\end{equation}

Equation \@ref(eq:bayes) seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

# Discussion

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional details

\newpage

# References

### 2022-03-19 outputs/paper/paper.Rmd
fvdvdv

@@
 author: 
-  - First author
-  - Another author
-thanks: "Code and data are available at: LINK."
+  - Adam Labas  
+  - Krupali Bhavsar
+thanks: "Code and data are available at: XXXXXXXXXXXXXXXXXXXXXXX"
 date: "`r format(Sys.time(), '%d %B %Y')`"

### 2022-03-19 scripts/01-data_cleaning.R
commiiiter

@@ -1,14 +1,15 @@
+comments
+Code 
+notes for Krupali and Adam
+
 #### Preamble ####
-# Purpose: Clean the survey data downloaded from [...UPDATE ME!!!!!]
-# Author: Rohan Alexander [CHANGE THIS TO YOUR NAME!!!!]
-# Data: 3 January 2021
-# Contact: rohan.alexander@utoronto.ca [PROBABLY CHANGE THIS ALSO!!!!]
-# License: MIT
+# Purpose: Prepare the 2021 GSS data
+# Author: Adam Labas and Krupali Bhavsar
+# Data: March 20 2022
+# Contact: adam.labas@mail.utoronto.ca and krupali.bhavsar@mail.utoronto.ca
+# License: 
 # Pre-requisites: 
-# - Need to have downloaded the ACS data and saved it to inputs/data
-# - Don't forget to gitignore it!
-# - Change these to yours
-# Any other information needed?
+# - Need to have downloaded the GSS data and saved it to inputs/data
 
 
 #### Workspace setup ####
@@ -16,21 +17,19 @@
 library(haven)
 library(tidyverse)
 # Read in the raw data. 
-raw_data <- readr::read_csv("inputs/data/raw_data.csv"
-                     )
+raw_data <- haven::read_dta("inputs/data/2021_stata/gss2021.dta")
+
 # Just keep some variables that may be of interest (change 
 # this depending on your interests)
 names(raw_data)
 
 reduced_data <- 
   raw_data %>% 
-  select(first_col, 
-         second_col)
+  select(age, 
+         sexnow1) over here we have to pic what variables/columns we want to keep
 rm(raw_data)
-         
+
 
 #### What's next? ####
-
-
 
          

### 2022-03-19 scripts/01-data_cleaning.R
cmittment issues

@@ -1,7 +1,3 @@
-comments
-Code 
-notes for Krupali and Adam
-
 #### Preamble ####
 # Purpose: Prepare the 2021 GSS data
 # Author: Adam Labas and Krupali Bhavsar
@@ -26,9 +22,15 @@ names(raw_data)
 reduced_data <- 
   raw_data %>% 
   select(age, 
-         sexnow1) over here we have to pic what variables/columns we want to keep
-rm(raw_data)
+         sexnow1,
+         born,
+         marital,
+         educ,
+         divorce,
+         degree)
 
+rm(raw_data)
+#over here we have to pic what variables/columns we want to keep
 
 #### What's next? ####
 

### 2022-03-19 scripts/01-data_cleaning.R
test test

@@ -1,4 +1,4 @@
-#### Preamble ####
+#### Preamble dksbcasvkfvkcssf####
 # Purpose: Prepare the 2021 GSS data
 # Author: Adam Labas and Krupali Bhavsar
 # Data: March 20 2022

### 2022-03-20 outputs/paper/paper.Rmd
second commit

--- 2022-03-19 outputs/paper/paper.Rmd
@@
 ---
-title: "My titleXXXXXXXXXXXXXXXX"
-subtitle: "My subtitle if needed"
+title: "Income Has Serious Implications On Many Facets of American Life in 2021"
+subtitle: "A Look Into The Implications of Americans' Incomes Reveals Information On Disparity of The Gender Pay and the American Dream."
 author: 
-  - First author
-  - Another author
-thanks: "Code and data are available at: LINK."
+  - Adam Labas  
+  - Krupali Bhavsar
+thanks: "Code and data are available at: XXXXXXXXXXXXXXXXXXXXXXX"
 date: "`r format(Sys.time(), '%d %B %Y')`"
@@
 
-```{r setup, include=FALSE}
+```{r setup, fig.pos = "!H", include = FALSE}
 knitr::opts_chunk$set(echo = TRUE)
@@
 library(tidyverse)
-library(palmerpenguins)
+library(janitor)
+library(dplyr)
+library(tidyr)
+library(opendatatoronto)
+library(knitr)
+library(patchwork)
+library(lubridate)
 ```
@@
 
-Our data is of penguins (Figure \@ref(fig:bills)).
+In this Data Section \@ref(data), we will provide a look into the data acquisition and processing methodology as well as a deep dive into the contents of the data.
+First of, (Table \@ref(tab:tableofdata)) give us a glimpse of the data.
 
-```{r bills, fig.cap="Bills of penguins", echo = FALSE}
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = FALSE) +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
-```
+## Data Collection
 
-Talk more about it.
-
-Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)
-
-```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=8, fig.height=4}
-
-# This needs to be about the random data tha tI sasmpled?
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = FALSE) +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
-```
+All of the data used in this analysis was directly collected from the United States 2021 General Social Survey, available here: https://gss.norc.org/ (link and additional information available in the README.md file.) We downloaded the desired data in the STATA format.
+The zip file acquired included 4 files: a read me text file, the raw data as a dta file, as well as two pdf files.
+*GSS 2021 Codebook R1b* is essentially the instruction manual for the data set.
+It does over
 

### 2022-03-20 scripts/01-data_cleaning.R

@@ -1,37 +1,0 @@
-#### Preamble dksbcasvkfvkcssf####
-# Purpose: Prepare the 2021 GSS data
-# Author: Adam Labas and Krupali Bhavsar
-# Data: March 20 2022
-# Contact: adam.labas@mail.utoronto.ca and krupali.bhavsar@mail.utoronto.ca
-# License: 
-# Pre-requisites: 
-# - Need to have downloaded the GSS data and saved it to inputs/data
-
-
-#### Workspace setup ####
-# Use R Projects, not setwd().
-library(haven)
-library(tidyverse)
-# Read in the raw data. 
-raw_data <- haven::read_dta("inputs/data/2021_stata/gss2021.dta")
-
-# Just keep some variables that may be of interest (change 
-# this depending on your interests)
-names(raw_data)
-
-reduced_data <- 
-  raw_data %>% 
-  select(age, 
-         sexnow1,
-         born,
-         marital,
-         educ,
-         divorce,
-         degree)
-
-rm(raw_data)
-#over here we have to pic what variables/columns we want to keep
-
-#### What's next? ####
-
-         

### 2022-03-20 scripts/Data Acquisition and Processing.R

@@ -1,0 +1,37 @@
+#### Preamble dksbcasvkfvkcssf####
+# Purpose: Prepare the 2021 GSS data
+# Author: Adam Labas and Krupali Bhavsar
+# Data: March 20 2022
+# Contact: adam.labas@mail.utoronto.ca and krupali.bhavsar@mail.utoronto.ca
+# License: 
+# Pre-requisites: 
+# - Need to have downloaded the GSS data and saved it to inputs/data
+
+
+#### Workspace setup ####
+# Use R Projects, not setwd().
+library(haven)
+library(tidyverse)
+# Read in the raw data. 
+raw_data <- haven::read_dta("inputs/data/2021_stata/gss2021.dta")
+
+# Just keep some variables that may be of interest (change 
+# this depending on your interests)
+names(raw_data)
+
+reduced_data <- 
+  raw_data %>% 
+  select(age, 
+         sexnow1,
+         born,
+         marital,
+         educ,
+         divorce,
+         degree)
+
+rm(raw_data)
+#over here we have to pic what variables/columns we want to keep
+
+#### What's next? ####
+
+         

### 2022-03-21 outputs/paper/paper.Rmd
third commit

@@
 knitr::opts_chunk$set(echo = TRUE)
-
+install.packages("tinytex")
+tinytex::install_tinytex()
+library(tinytex)
 library(tidyverse)
@@
 library(tidyr)
-library(opendatatoronto)
 library(knitr)
@@
 
-All of the data used in this analysis was directly collected from the United States 2021 General Social Survey, available here: https://gss.norc.org/ (link and additional information available in the README.md file.) We downloaded the desired data in the STATA format.
+All of the data used in this analysis was directly collected from the United States 2021 General Social Survey, available here: https://gss.norc.org/ (link and additional information available in the README.md file.) We downloaded the desired data in the STATA format and loaded it into R [@citeR] for manipulation.
 The zip file acquired included 4 files: a read me text file, the raw data as a dta file, as well as two pdf files.
 *GSS 2021 Codebook R1b* is essentially the instruction manual for the data set.
-It does over
+It goes over many topics like the objectives and intents of the survey, a brief overview of the methodology and an in depth explanation of the 565 variables in the raw data.
+With 565 variables collected in the survey, the data file titles *gss2021.dta* has a size of 2.6 MB.
+Although this isn't too large of a data file, we did encounter some problems which may have implications as it pertains to reproducibility of the workflow.
+The following is an explanation of the problems encountered as well as an explanation of the workaround/solution we implemented to get over the obstacle.
 
-Talk way more about it.
+## Problem With Data Upload into R Studio
+We know that having a reproducible workflow is crucial if we want to produce mature and ethical statistical papers which any data scientist down the line can reproduce for any purpose from verifying results or modifying methodology.
+As previously stated, the raw data provided by the United States 2021 General Social Survey was in a zip folder and in dta format.
+To conduct the development of this paper, we worked in RStudio {@citeR) in a server developed, maintained an administered by one of the TAs for the STA304 Course at the University of Toronto (Winter 2022), Mauricio Vargas Sepúlveda, better known as Pachá.
+This server works just like any other server for RStudio like JupyterHub but the admin user, Pachá, has the ability to overview all activity on the server and access users accounts to aid with trouble shooting or data sharing: this will proved to be decisive in just a moment.
+When we tried to follow typical procedure to upload the 2021_stata.zip file into under data in the inputs folder, we received a message we had never received before.
+Figure 1 is a screen capture of that error.
+![The Error received upon attempting to upload the zip folder](~/paper_3_team_paper/inputs/2021_stata_error_message.png "The error message we received when trying to upload the zip folder.")
 
-# Model
+This was hard for us to wrap our heads around as we were able to upload other files independently, and crucially we were able to upload the other file in the zip independently aside from the dta file.
+We tried to replicate this problem working in RStudio locally but surprisingly it was not a problem there.
+As such, we contacted the server admin, shared the zip folder with him, and using his administrator permissions, he uploaded the unzipped folder into a shared folder that the entire class has access to.
+We then copped this folder into the data folder under inputs.
+It is important to note that throughout this process, none of the information from the zip folder had been altered or manipulated and that all the information, and only the information, which had started in it, ended up in our instance of RStudio.
 
-\begin{equation}
-Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}  (\#eq:bayes)
-\end{equation}
+Lastly on this matter, should you run into this problem while attempting to reproduce our findings or work, working locally in RStudio might fix the problem.
 
-Equation \@ref(eq:bayes) seems useful, eh?
+## Data Processing and Variable Selection
 
-Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR].
-We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.
+Upon being able to successfully load the data into R (@citeR) we began data processing and variable selction.
+This includes things like cleaning the data from missing values, choosing the variables of interest and producing new variables which will help us with our analysis.
+The following is an overview of the methods and events which took place in this phase of the analysis.
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+All the data processing, cleaning and variable selection takes place in an R script called *Data Acquisition and Processing.R* in the scripts folder.
+To process our data, we began by loading all the necessary libraries like haven [@citehaven] which was used to acess the information from the dta type file and we then load the data into R [@citeR].
+Additionally, we used a wide variety of libraries like knitr [@citeknitr], tidyverse [@citetidyverse], tidyr [@citetidyr], janitor [@citejanitor], dplyr [@citetidyr] as well as tinytex [@citetinytex] at some points in the process for data processing and pdf document generation.
+
+After loading the data into our local environment and saving its as a cvs called raw.csv, we began to create as subset of the data with the desired variables.
+There are 565 variables in the data and we obviously can't explain each of them, so here is a description of some of the variables we chose to work with.
+*sexnow1* is the current sex of the respondent and is a concept more similar to gender, *age*, *born*: born in America or not, *marital*: the marital status, *educ*: the highest attained level of education, *divorce*: divroced or not, whether or not they have a *degree*, *income*: their current family income, *income16*: their family income when they were 16 years old, *nathealy*: if they think enough is done for healthcare, *cohort*: year of birth, *agekdbrn*: age when first child was born and lastly, *granborn* which is a number from 0-4 signifying the number of grandparents of the respondent which were born in the united states.
+
+Additionally, we created a new column called *income_cat* and *income16_cat* which are new variables which offer us the liberty of using the *income* and *income16* variables but as discreet, categorical variables instead of the coded numerical system used all throughout the data set and explained in the information code book titled *GSS 2021 Codebook R1b.pdf* available in the 2021_stata in the data folder in inputs.This was done by duplicating the original columns and then re-coding the values.
+In the case of *income* for instance, the codebook explains that a value of 1 for income means the response was that the annual income was less than \$1000.
+As such, income_cat would read *$1,000 AND UNDER*.
+Similar mutations were done for the variables *sex*, *granborn*, *nathealy* and *born*.
+
+We finished off by saving all the changed in a new data set: reduced_data.csv
+
+# Methodology
+
+In matters of reproducibility, the methodology used to collect data is always important.
+This includes the various sampling techniques that were used, the way that non responses are dealt with and other key decisions taken by the data collection team and the survey itself.
+The following is an explanation of the methodology used in the 2021 GSS, as well as its key features, strengths, and weaknesses.
+
+## Methodology Overview
+
+The 2021 GSS is in many ways different to the previous General Social Surveys in years prior to the ascension of the corona virus to international stardom.
+According to the GSS 2021 Codebook R1b, in previous iterations of the GSS, the survey was taken solely via in-person interview, where the interviewer had been adequately trained and prepared for various situations that would have arose like non responds or the person being interviewed not knowing an answer.
+However, for reasons related to health and safety protocols and "to safeguard the health of staff and respondents during the COVID-19 pandemic, the 2021 GSS data collection used a mail-to-web methodology instead of its traditional in-person interviews." This means that after being selected via address based sampling the selected individuals "interview" would be conducted online via a self administered questionnaire.
+
+## Key features
 

### 2022-03-21 scripts/Data Acquisition and Processing.R

@@ -12,12 +12,21 @@
 # Use R Projects, not setwd().
 library(haven)
 library(tidyverse)
+library(janitor)
+library(tidyr)
+library(dplyr)
+library(knitr)
+library(patchwork)
 # Read in the raw data. 
 raw_data <- haven::read_dta("inputs/data/2021_stata/gss2021.dta")
+
+write.csv(x = raw_data,
+          file = "inputs/data/raw.csv")
 
 # Just keep some variables that may be of interest (change 
 # this depending on your interests)
 names(raw_data)
+
 
 reduced_data <- 
   raw_data %>% 
@@ -27,11 +36,100 @@ reduced_data <-
          marital,
          educ,
          divorce,
-         degree)
+         degree,
+         income,
+         nathealy,
+         cohort,
+         agekdbrn,
+         income16,
+         granborn) %>% 
+  rename(sex = sexnow1)
 
-rm(raw_data)
-#over here we have to pic what variables/columns we want to keep
+reduced_data$income16_cat <- reduced_data$income16
+reduced_data$income_cat <- reduced_data$income
 
-#### What's next? ####
+reduced_data <- 
+  reduced_data %>%
+  mutate(sex = case_when(
+    sex == 1 ~ "Male",
+    sex == 2 ~ "Female",
+    sex == 3 ~ "Transgender",
+    sex == 4 ~ "None of These"
+  ))
 
-         
+reduced_data <- 
+  reduced_data %>% 
+  mutate(granborn = case_when(
+    granborn == 0 ~ "None",
+    granborn == 1 ~ "One",
+    granborn == 2 ~ "Two",
+    granborn == 3 ~ "Three",
+    granborn == 4 ~ "Four"
+  ))
+
+reduced_data <- 
+  reduced_data %>% 
+  mutate(born = case_when(
+    born == 1 ~ "Yes",
+    born == 2 ~ "No"
+  ))
+
+reduced_data <- 
+  reduced_data %>%
+  mutate(nathealy = case_when(
+    nathealy == 1 ~ "Too Little",
+    nathealy == 2 ~ "About Right",
+    nathealy == 3 ~ "Too Much"
+  ))
+
+reduced_data <- 
+  reduced_data %>% 
[…]

### 2022-03-21 outputs/paper/paper.Rmd
fourth commit

@@
 title: "Income Has Serious Implications On Many Facets of American Life in 2021"
-subtitle: "A Look Into The Implications of Americans' Incomes Reveals Information On Disparity of The Gender Pay and the American Dream."
+subtitle: "A Look Into The Implications of Americans' Incomes Reveals Information On The Veracity of American Dream."
 author: 
@@
   bookdown::pdf_document2
-toc: FALSE
+toc: TRUE
 bibliography: references.bib
@@
 knitr::opts_chunk$set(echo = TRUE)
-install.packages("tinytex")
-tinytex::install_tinytex()
+# install.packages("tinytex")
+# tinytex::install_tinytex()
 library(tinytex)
@@
 library(lubridate)
+
+reduced_data <- read_csv(here::here("inputs/data/reduced_data.csv"))
+
 ```
@@
 First of, (Table \@ref(tab:tableofdata)) give us a glimpse of the data.
+
+```{r echo = F}
+#table of variables that I am concerned with
+
+reduced_data[complete.cases(reduced_data), ] %>%
+   slice(1:7) %>%
+  kable(
+    caption = "GSS Cleaned Data",
+    col.names = c("Row Number", "Age", "Gender", "Born", "Marital Status", "Education", 
+                  "Divorce Status", "Degree", "Current Income", "Health Care", "Birth Year", 
+                  "First Born", "Income at 16", "Grandparents", "Category for Income at 16", 
+                  "Caegory for Current Income"),
+    digits = 1,
+    booktabs = TRUE,  
+    linesep = "")
+```
 
@@
 Additionally, we created a new column called *income_cat* and *income16_cat* which are new variables which offer us the liberty of using the *income* and *income16* variables but as discreet, categorical variables instead of the coded numerical system used all throughout the data set and explained in the information code book titled *GSS 2021 Codebook R1b.pdf* available in the 2021_stata in the data folder in inputs.This was done by duplicating the original columns and then re-coding the values.
-In the case of *income* for instance, the codebook explains that a value of 1 for income means the response was that the annual income was less than \$1000.
+In the case of *income* for instance, the code book explains that a value of 1 for income means the response was that the annual income was less than \$1000.
 As such, income_cat would read *$1,000 AND UNDER*.
@@
 However, for reasons related to health and safety protocols and "to safeguard the health of staff and respondents during the COVID-19 pandemic, the 2021 GSS data collection used a mail-to-web methodology instead of its traditional in-person interviews." This means that after being selected via address based sampling the selected individuals "interview" would be conducted online via a self administered questionnaire.
+The limitations of such will later be discussed in Section \@ref(limitations)
 
@@
 
+It is important to discuss the reach of our data.
+We know from the provided documentation that the population group eligible for taking the survey are Adults 18 or older in the United States who live in non institutional housing at the time of interviewing.
+Also, the 2021 GSS  uses the last birthday method to decide which individual in the household will take the survey.
+This was done by having "a professional phone interviewer team conducted phone outreach to complete screeners, answer inbound phone calls from sampled households, identify respondents, and complete interviews by phone and/or prompt them to complete the web survey." As far as the invitation method goes for data collection, materials that provide a web link to invite people to participate in the survey was mailed out and an option to conduct the interview via the phone instead of a self-administered questionnaire was also provided.
+Out of the 27,591 lines of sample (people who had been prompted to answer the questionnaire), 4,032 responses were collected.
+This has implications on the finding of our analysis in Section \@ref(results) as we will have to consider if we are able to generalize our finding from the relatively small sample that we have to the population if all US citizens over 18 years old.
+The small sample size and extremely long survey are both weaknesses of the survey which we will discuss in more depth in Section \@ref(results).
+
+In addition, the 2021 GSS used census data as well as publicly available information to try and identify areas and specific household with Spanish speakers whereby the survey was then offered in English and in Spanish.
+This is a big strength of the survey at it shows the attempt of the data analyst to be as mature and ethical as possible and offer the survey in a language that is more suitable for a large portion of US citizens.
+
 # Results
+
+```{r}
+#creating levels for the x and y axis on the graphs
+
+level_income16 <- c("$1,000 AND UNDER","$1,000 TO $2,999","$3,000 TO $3,999",
+                    "$4,000 TO $4,999","$5,000 TO $5,999","$6,000 TO $6,999",
+                    "$7,000 TO $7,999","$8,000 TO $9,999","$10,000 TO $12,499",
+                    "$12,500 TO $14,999", "$15,000 TO $17,499","$17,500 TO $19,999",
+                    "$20,000 TO $22,499","$22,500 TO $24,999","$25,000 TO $29,999",
+                    "$30,000 TO $34,999","$35,000 TO $39,999","$40,000 TO $49,999",
+                    "$50,000 TO $59,999","$60,000 TO $74,999","$75,000 TO $89,999",
+                    "$90,000 TO $109,999","$110,000 TO $129,999","$130,000 TO $149,999",
+                    "$150,000 TO $169,999","$170,000 OR OVER")
+
+level_income <- c("$1,000 AND UNDER","$1,000 TO $2,999","$3,000 TO $3,999",
+                    "$4,000 TO $4,999","$5,000 TO $5,999","$6,000 TO $6,999",
+                    "$7,000 TO $7,999","$8,000 TO $9,999","$10,000 TO $14,999",
+                    "$15,000 TO $19,999","$20,000 TO $24,999","$25,000 OR MORE")
+```
+
+```{r fig1, fig.cap="write a figure caption here", echo = FALSE, fig.width=12, fig.height=10, echo=FALSE}
+#income16 vs income grouped by gender
+
+income16_gender <- ggplot(mapping = aes(x = factor(income16_cat, level = level_income16), group = sex, fill = sex), data = reduced_data) +
+  geom_bar(alpha = 0.7)  +
+  theme_minimal() +
+  labs(x = "Family Income at 16 Years Old",
+       y = "Number of occurrences",
+       fill = "Gender",
+       title = "Total Family Income at 16 Grouped by Gender") +
+  coord_flip()
+
+income_gender <- ggplot(mapping = aes(x = factor(income_cat, level = level_income), group = sex, fill = sex), data = reduced_data) +
+  geom_bar(alpha = 0.7)  +
+  theme_minimal() +
+  labs(x = "Family Income",
+       y = "Number of occurrences",
+       fill = "Gender",
+       title = "Total Family Income Grouped by Gender") +
+  coord_flip()
+
+income16_gender/income_gender
+```
+
+```{r, fig.cap="write a figure caption here", echo = FALSE, fig.width=12, fig.height=10, echo=FALSE}
+#fincome16 vs income grouped by where grandparents were born
+
+income16_grand <- ggplot(mapping = aes(x = factor(income16_cat, level = level_income16), group = granborn, fill = granborn), data = reduced_data) +
+  geom_bar(alpha = 0.7)  +
+  theme_minimal() +
+  labs(x = "Family Income at 16 Years Old",
+       y = "Number of occurrences",
+       fill = "Grandparents",
+       title = "Total Family Income at 16 Grouped by Grandparent Birthplace") +
+  coord_flip()
+
+income_grand <- ggplot(mapping = aes(x = factor(income_cat, level = level_income), group = granborn, fill = granborn), data = reduced_data) +
+  geom_bar(alpha = 0.7)  +
+  theme_minimal() +
+  labs(x = "Family Income",
+       y = "Number of occurrences",
+       fill = "Grandparents",
+       title = "Total Family Income Grouped by Grandparent Birthplace") +
+  coord_flip()
+
+income16_grand/income_grand
+```
+
+```{r, fig.cap="write a figure caption here", echo = FALSE, fig.width=12, fig.height=10, echo=FALSE}
+#family income at 16 grouped by where individual was born
+
+income16_born <- ggplot(mapping = aes(x = factor(income16_cat, level = level_income16), group = born, fill = born), data = reduced_data) +
+  geom_bar(alpha = 0.7)  +
+  theme_minimal() +
+  labs(x = "Total Family Income at 16 Years Old",
+       y = "Number of occurrences",
+       fill = "Born",
+       title = "Total Family Income at 16 Grouped by individual’s Birthplace") +
+  coord_flip()
+
+income_born <- ggplot(mapping = aes(x = factor(income_cat, level = level_income), group = born, fill = born), data = reduced_data) +
+  geom_bar(alpha = 0.7)  +
+  theme_minimal() +
+  labs(x = "Total Family Income",
+       y = "Number of occurrences",
+       fill = "Born",
+       title = "Total Family Income Grouped by individual’s Birthplace") +
+  coord_flip()
+
+income16_born/income_born
+```
+
+```{r, fig.cap="write a figure caption here", echo = FALSE, fig.width=9, fig.height=6, echo=FALSE}
+#income at 16 vs current income grouped by where individual was born
+
+income_income16_born <- ggplot(mapping = aes(x = income_cat, y = income16_cat, color = born), data = reduced_data) +
+  geom_point() +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 90)) +
+  labs(x = "Current Family Income",
+       y = "Family Income at Age 16",
+       color = "Born in The USA",
+       title = "Relationship Between Family Income at age 16 and Current Family Income")
+
+income_income16_born
+```
 

### 2022-03-21 outputs/paper/paper.Rmd
fifth commit

@@
 
-## Key features
+## Key features & Pros And Cons of the Methodology and Survey
 
@@
 
-# Results
-
-```{r}
+# Visualizing the Data and The Implications
+In this analysis, I am interested at looking at how real is the American Dream and if it is found to be possible, whether or not there is any difference in attainment amoung genders.
+As such, in this Section \@ref(visualizingthedataandtheimplications), we will create various visuals which will enable us to extract as much information as we can and we will speak about the implications that they make.
+```{r, echo = F, message = F, include = F}
 #creating levels for the x and y axis on the graphs
@@
 
-```{r fig1, fig.cap="write a figure caption here", echo = FALSE, fig.width=12, fig.height=10, echo=FALSE}
+## Visual 1
+```{r figure-1, fig.cap="write a figure caption here", echo = FALSE, fig.width=12, fig.height=10, message = F, warning = F}
+income16_gender <- ggplot(mapping = aes(x = factor(income16_cat, level = level_income16), group = sex, fill = sex), data = reduced_data) +
+  geom_bar(alpha = 0.7)  +
+  theme_minimal() +
+  labs(x = "Family Income at 16 Years Old",
+       y = "Number of occurrences",
+       fill = "Gender",
+       title = "Total Family Income at 16 Grouped by Gender") +
+  coord_flip()
+income16_gender
+```
+In this graph (Figure \@ref(fig:figure-1)), we can proportions of family income when the responded was 16 years of age.
+We also see that an over-whelming majority of people in our sample had an income of over $40,000.
+We also see that the single largest individual group is the *NA* group.
+We will discuss the implications of this in Section \@ref(limitations1).
+As we've previously discussed, when the questioned is prompted with a question, in every situation, they have the option to refuse to answer the question or say that they do not know.
+The both of these are listed as NA responses.
+We also see that of the 5 possible genders (including the NA gender, not to be confused with the NA response to income16), the overwhelming majority of respondents identified as male or female as expected.
+
+We would like to contrast this with the current income of responds. The below is the same graph above but we've also included a similar graph but for family income at the time of response 
+```{r figure-2, fig.cap="write a figure caption here", echo = FALSE, fig.width=12, fig.height=10, message = F, warning = F}
 #income16 vs income grouped by gender
@@
 ```
+In figure \@ref(fig:figure-2), we see the overlapping of the two graphs, the first of which is the same as figure \@ref(fig:fig1) and the bottom one is the same in essence but with the information for current family income instead of at age 16.
+The first and most striking part of this visual is that for the current income, the highest thres
 
-```{r, fig.cap="write a figure caption here", echo = FALSE, fig.width=12, fig.height=10, echo=FALSE}
+```{r, fig.cap="write a figure caption here", echo = FALSE, fig.width=12, fig.height=10, message = F, warning = F}
 #fincome16 vs income grouped by where grandparents were born
@@
 
-```{r, fig.cap="write a figure caption here", echo = FALSE, fig.width=12, fig.height=10, echo=FALSE}
+```{r, fig.cap="write a figure caption here", echo = FALSE, fig.width=12, fig.height=10, message = F, warning = F}
 #family income at 16 grouped by where individual was born
@@
 
-```{r, fig.cap="write a figure caption here", echo = FALSE, fig.width=9, fig.height=6, echo=FALSE}
+```{r, fig.cap="write a figure caption here", echo = FALSE, fig.width=9, fig.height=6, message = F, warning = F}
 #income at 16 vs current income grouped by where individual was born

### 2022-03-21 outputs/paper/paper.Rmd
fifth commit

@@
 Out of the 27,591 lines of sample (people who had been prompted to answer the questionnaire), 4,032 responses were collected.
-This has implications on the finding of our analysis in Section \@ref(results) as we will have to consider if we are able to generalize our finding from the relatively small sample that we have to the population if all US citizens over 18 years old.
+This has implications on the finding of our analysis in Section \@ref(discussion) as we will have to consider if we are able to generalize our finding from the relatively small sample that we have to the population if all US citizens over 18 years old.
 The small sample size and extremely long survey are both weaknesses of the survey which we will discuss in more depth in Section \@ref(results).
@@
 In this analysis, I am interested at looking at how real is the American Dream and if it is found to be possible, whether or not there is any difference in attainment amoung genders.
-As such, in this Section \@ref(visualizingthedataandtheimplications), we will create various visuals which will enable us to extract as much information as we can and we will speak about the implications that they make.
+As such, in this Section \@ref(visualizing-the-data-and-the-implications), we will create various visuals which will enable us to extract as much information as we can and we will speak about the implications that they make.
 ```{r, echo = F, message = F, include = F}
@@
 ## Visual 1
-```{r figure-1, fig.cap="write a figure caption here", echo = FALSE, fig.width=12, fig.height=10, message = F, warning = F}
+```{r figure 1, fig.cap="write a figure caption here", echo = FALSE, fig.width=12, fig.height=10, message = F, warning = F}
 income16_gender <- ggplot(mapping = aes(x = factor(income16_cat, level = level_income16), group = sex, fill = sex), data = reduced_data) +
@@
 We also see that the single largest individual group is the *NA* group.
-We will discuss the implications of this in Section \@ref(limitations1).
+We will discuss the implications of this in Section \@ref(limitations).
 As we've previously discussed, when the questioned is prompted with a question, in every situation, they have the option to refuse to answer the question or say that they do not know.
@@
 We would like to contrast this with the current income of responds. The below is the same graph above but we've also included a similar graph but for family income at the time of response 
-```{r figure-2, fig.cap="write a figure caption here", echo = FALSE, fig.width=12, fig.height=10, message = F, warning = F}
+```{r figure2, fig.cap="write a figure caption here", echo = FALSE, fig.width=12, fig.height=10, message = F, warning = F}
 #income16 vs income grouped by gender
@@
 ```
-In figure \@ref(fig:figure-2), we see the overlapping of the two graphs, the first of which is the same as figure \@ref(fig:fig1) and the bottom one is the same in essence but with the information for current family income instead of at age 16.
-The first and most striking part of this visual is that for the current income, the highest thres
+In figure \@ref(fig:figure-2), we see the overlapping of the two graphs, the first of which is the same as figure \@ref(fig:figure-1) and the bottom one is the same in essence but with the information for current family income instead of at age 16.
+The first and most striking part of this visual is that for the current income, the highest threshold is $25,000 or more.
+We see that the both of our graphs are heavily skewed to wards higher incomes.
+I suspect that this is mainly a production of the bin widths decided on for the analysis.
+
+## Implications 1
+In Visuals \@ref(visual-1) we can see when we look into income16 that the majority of people claimed that at the age of 16, their family income was over 40 thousand dollars.
+When we lookover at the income variable, we are trully perplexed to see that the largest is 25 thousand or more.
+We will discuss the implications of the latter in Section \@ref(limitations).
+We can also see that the number of NA responses rises from about 530 in the case of the income at age 16 variable to approximately 900 in the income variable.
+This seems counter intuitive as it would appear that less people are uncertain about their family income at age 16 than they are.
+We suspect however that this is not true.
+We suspect that there could be a possible shame or embarassment assosociated with sharing ones current faily income as it represents the persons way of living or lifestyle today.
+On the other hand, someones incomeat age 16 is psychologically less of importance to them and carries no burden in their life, as it is notan expected metric of succes year down the line.
+We also recognize that amongst men and women, there is a general
 
@@
 
-# Discussion
+# Limitations
 
 ## First discussion point
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
 
@@
 
-## Weaknesses and next steps
+## Next steps
 

### 2022-03-21 outputs/paper/paper.Rmd
6th commit

@@
   - Krupali Bhavsar
-thanks: "Code and data are available at: XXXXXXXXXXXXXXXXXXXXXXX"
+thanks: "Code and data are available at: https://github.com/adam-labas/The-US-GSS-And-The-American-Dream"
 date: "`r format(Sys.time(), '%d %B %Y')`"
@@
 ![The Error received upon attempting to upload the zip folder](~/paper_3_team_paper/inputs/2021_stata_error_message.png "The error message we received when trying to upload the zip folder.")
+
+{ width="800" height="600" style="display: block; margin: 0 auto" }
 
@@
 
-We would like to contrast this with the current income of responds. The below is the same graph above but we've also included a similar graph but for family income at the time of response 
+We would like to contrast this with the current income of responds. The below is the same graph above but we've also included a similar graph but for family income at the time of response.
 ```{r figure2, fig.cap="write a figure caption here", echo = FALSE, fig.width=12, fig.height=10, message = F, warning = F}
@@
 In Visuals \@ref(visual-1) we can see when we look into income16 that the majority of people claimed that at the age of 16, their family income was over 40 thousand dollars.
-When we lookover at the income variable, we are trully perplexed to see that the largest is 25 thousand or more.
+When we look over at the income variable, we are truly perplexed to see that the largest is 25 thousand or more.
 We will discuss the implications of the latter in Section \@ref(limitations).
@@
 We suspect however that this is not true.
-We suspect that there could be a possible shame or embarassment assosociated with sharing ones current faily income as it represents the persons way of living or lifestyle today.
-On the other hand, someones incomeat age 16 is psychologically less of importance to them and carries no burden in their life, as it is notan expected metric of succes year down the line.
-We also recognize that amongst men and women, there is a general
+We suspect that there could be a possible shame or embarrassment associated with sharing ones current family income as it represents the persons way of living or lifestyle today.
+When we speak of the American dream, we implicitly assume an increase in wealth over time and for someone who isn't as wealthy as they might want to be (likely an overwhelming large portion of the population), reluctance with sharing their current income is totally understandable.
+On the other hand, someones income at age 16 is psychologically less of importance to them and carries no burden in their life, as it is not an expected metric of success years down the line.
+We also recognize that among men and women, there is a general almost a 50:50 split at each level of income but we do notice that for income level NA, a lot of people chose not to identify their gender.
+It would be interesting in another analysis to analyse possible evidences of gender pay gap and how people who do not fit in the *traditional* gender roles are affected by the advent of the American Dream and whether or not their position in life hinders their ability to increse in wealth over time.
+
+As we speak about an increase in wealth over time, we next want to see if the birth place of ones grandparents has any impact on income due to the so called *generational wealth*.
+
+## Visual 2
+As weve previously stated, the American
+\@ref(visual-2)
 
@@
 
-```{r, fig.cap="write a figure caption here", echo = FALSE, fig.width=9, fig.height=6, message = F, warning = F}
-#income at 16 vs current income grouped by where individual was born
-
-income_income16_born <- ggplot(mapping = aes(x = income_cat, y = income16_cat, color = born), data = reduced_data) +
-  geom_point() +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90)) +
-  labs(x = "Current Family Income",
-       y = "Family Income at Age 16",
-       color = "Born in The USA",
-       title = "Relationship Between Family Income at age 16 and Current Family Income")
-
-income_income16_born
-```
-
 # Limitations
@@
 
-# Additional details
-
 \newpage

### 2022-03-21 outputs/paper/paper.Rmd
7th commit

@@
 ## Visual 2
-As weve previously stated, the American
-\@ref(visual-2)
-
+As we have previously stated, the American dream is one which communicates the increase of wealth over time for hard working Americans, especially to the eyes of immigrants.
+As such, we will attempt to visualize the impact of grandparent immigration to the United States on the income of Americans.
 ```{r, fig.cap="write a figure caption here", echo = FALSE, fig.width=12, fig.height=10, message = F, warning = F}
@@
 
-```{r, fig.cap="write a figure caption here", echo = FALSE, fig.width=12, fig.height=10, message = F, warning = F}
+## Implications 2
+In the above visuals, we can see that in all the income groupings for both age 16 and current income, the largest subgrouping within each is people who have no grandparents born in the United states.
+The second largest group are usually those with 2 grandparents born in the United States.
+It is pretty logical that 2 grandparents follows after zero grandparents as it is easy to recognize that most people do not go abroad to marry and marry a friend or someone that they know locally, who is most likely to be born in the united states.
+We would speculate that a similar trend can be observed in other *immigrant* countries Like Canada or Sweeden.
+What is the most fascinating however is that the vast majority of people who answered the survey are first or second generation Americans, perhaps having grandparents or parents who immigrated to this country chasing the American Dream.
+
+In a similar fashion to what was explained in \@ref(implications-1), both of our graphs are heavily skewed towards higher income.
+In the case of current income, it is almost certainly because of the bin width decided on by the people who built the survey which we personally find many problems with.
+In Section \@ref(limitations), we will take a deep dive into this.
+
+## Visuals 3
+```{r echo=FALSE, fig.cap="write a figure caption here", fig.height=10, fig.width=12, message=FALSE, warning=FALSE}
 #family income at 16 grouped by where individual was born
@@
 # Limitations
+In this limitations section, I want to focus on the data which is provided to us and have a  look into why it might not be easy to take the implications mentioned above at face value
 
-## First discussion point
+## Uneven Bin width for Income Groups.
+The first major flaw I've found in the data is that the bins that both the income at age 16 and current income are recorded in are not all the same size.
+For example, in the case of the the first bin has a size of 16 years, the second bin from ages 16 to 24 has width 9, the next two have width 19 with the last ranging in theory from 65 to infinity.
+Although it is possible to understand why the maintainers of the data choose to have a demarcation point of the bins at 65 and 16, it is very hard to understand why they opted to divide the interior portion into varying bin sizes.
 
-## Second discussion point
+We know in data science that when plotting bar charts, the bin width can have a big implication on the presentation of the data and the things that we can infer about its distribution.
+If the bin width is too small or too large, we might reduce or add characteristics to the data that aren't actually representative of it.
 
+## Why cut off income at $25,000?
+use the us labor statistics data
 ## Third discussion point

### 2022-03-21 outputs/paper/paper.Rmd
8th commit

@@
 It goes over many topics like the objectives and intents of the survey, a brief overview of the methodology and an in depth explanation of the 565 variables in the raw data.
-With 565 variables collected in the survey, the data file titles *gss2021.dta* has a size of 2.6 MB.
+With 565 variables collected in the survey, the data file titled *gss2021.dta* has a size of 2.6 MB.
 Although this isn't too large of a data file, we did encounter some problems which may have implications as it pertains to reproducibility of the workflow.
@@
 ![The Error received upon attempting to upload the zip folder](~/paper_3_team_paper/inputs/2021_stata_error_message.png "The error message we received when trying to upload the zip folder.")
-
-{ width="800" height="600" style="display: block; margin: 0 auto" }
 
@@
 To process our data, we began by loading all the necessary libraries like haven [@citehaven] which was used to acess the information from the dta type file and we then load the data into R [@citeR].
-Additionally, we used a wide variety of libraries like knitr [@citeknitr], tidyverse [@citetidyverse], tidyr [@citetidyr], janitor [@citejanitor], dplyr [@citetidyr] as well as tinytex [@citetinytex] at some points in the process for data processing and pdf document generation.
+Additionally, we used a wide variety of libraries like knitr [@citeknitr], tidyverse [@citetidyverse], tidyr [@citetidyr], janitor [@citejanitor], patchwork [@citepatchwork], readr [@citereadr],dplyr [@citetidyr] as well as tinytex [@citetinytex] at some points in the process for data processing and pdf document generation.
 
-After loading the data into our local environment and saving its as a cvs called raw.csv, we began to create as subset of the data with the desired variables.
+After loading the data into our local environment and saving its as a cvs called raw.csv, we began to create subsets of the data with the desired variables.
 There are 565 variables in the data and we obviously can't explain each of them, so here is a description of some of the variables we chose to work with.
@@
 
-Additionally, we created a new column called *income_cat* and *income16_cat* which are new variables which offer us the liberty of using the *income* and *income16* variables but as discreet, categorical variables instead of the coded numerical system used all throughout the data set and explained in the information code book titled *GSS 2021 Codebook R1b.pdf* available in the 2021_stata in the data folder in inputs.This was done by duplicating the original columns and then re-coding the values.
+Additionally, we created new columns called *income_cat* and *income16_cat* which are new variables which offer us the liberty of using the *income* and *income16* variables as discreet, categorical variables instead of the coded numerical system used all throughout the data set and explained in the information code book titled *GSS 2021 Codebook R1b.pdf* available in the 2021_stata in the data folder in inputs.
+This was done by duplicating the original columns and then re-coding the values.
 In the case of *income* for instance, the code book explains that a value of 1 for income means the response was that the annual income was less than \$1000.
@@
 The 2021 GSS is in many ways different to the previous General Social Surveys in years prior to the ascension of the corona virus to international stardom.
-According to the GSS 2021 Codebook R1b, in previous iterations of the GSS, the survey was taken solely via in-person interview, where the interviewer had been adequately trained and prepared for various situations that would have arose like non responds or the person being interviewed not knowing an answer.
-However, for reasons related to health and safety protocols and "to safeguard the health of staff and respondents during the COVID-19 pandemic, the 2021 GSS data collection used a mail-to-web methodology instead of its traditional in-person interviews." This means that after being selected via address based sampling the selected individuals "interview" would be conducted online via a self administered questionnaire.
+According to the GSS 2021 Codebook R1b, in previous iterations of the GSS, the survey was taken solely via in-person interview, where the interviewer had been adequately trained and prepared for various situations that would have arose like non responses or the person being interviewed not knowing an answer.
+However, for reasons related to health and safety protocols and "to safeguard the health of staff and respondents during the COVID-19 pandemic, the 2021 GSS data collection used a mail-to-web methodology instead of its traditional in-person interviews." This means that after being selected via address based sampling the selected individual's "interview" would be conducted online via a self administered questionnaire.
 The limitations of such will later be discussed in Section \@ref(limitations)
@@
 It is important to discuss the reach of our data.
-We know from the provided documentation that the population group eligible for taking the survey are Adults 18 or older in the United States who live in non institutional housing at the time of interviewing.
+We know from the provided documentation that the population group eligible for taking the survey are adults 18 or older in the United States who live in non institutional housing at the time of interviewing.
 Also, the 2021 GSS  uses the last birthday method to decide which individual in the household will take the survey.
@@
 Out of the 27,591 lines of sample (people who had been prompted to answer the questionnaire), 4,032 responses were collected.
-This has implications on the finding of our analysis in Section \@ref(discussion) as we will have to consider if we are able to generalize our finding from the relatively small sample that we have to the population if all US citizens over 18 years old.
-The small sample size and extremely long survey are both weaknesses of the survey which we will discuss in more depth in Section \@ref(results).
+This has implications on the finding of our analysis in Section \@ref(limitations) as we will have to consider if we are able to generalize our finding from the relatively small sample that we have to the population of all US citizens over 18 years old.
+The small sample size and extremely long survey are both weaknesses of the survey which we will discuss in more depth in Section \@ref(limitations).
 
@@
 # Visualizing the Data and The Implications
-In this analysis, I am interested at looking at how real is the American Dream and if it is found to be possible, whether or not there is any difference in attainment amoung genders.
+In this analysis, I am interested at looking at how real the American Dream is and if it is found to be possible, whether or not there is any difference in attainment amoung genders.
 As such, in this Section \@ref(visualizing-the-data-and-the-implications), we will create various visuals which will enable us to extract as much information as we can and we will speak about the implications that they make.
@@
 ```
-In this graph (Figure \@ref(fig:figure-1)), we can proportions of family income when the responded was 16 years of age.
+In this graph (Figure \@ref(fig:figure-1)), we can see the proportions of family income when the responded was 16 years of age.
 We also see that an over-whelming majority of people in our sample had an income of over $40,000.
@@
 
-We would like to contrast this with the current income of responds. The below is the same graph above but we've also included a similar graph but for family income at the time of response.
+We would like to contrast this with the current income of responds. Below is the same graph as above, but we've also included a similar graph  for family income at the time of response.
 ```{r figure2, fig.cap="write a figure caption here", echo = FALSE, fig.width=12, fig.height=10, message = F, warning = F}
@@
 ## Implications 1
-In Visuals \@ref(visual-1) we can see when we look into income16 that the majority of people claimed that at the age of 16, their family income was over 40 thousand dollars.
-When we look over at the income variable, we are truly perplexed to see that the largest is 25 thousand or more.
+In Visuals \@ref(visual-1) we can see that when we look into income16  the majority of people claimed that at the age of 16, their family income was over 40 thousand dollars.
+When we look over at the income variable, we are truly perplexed to see that the largest bin is 25 thousand or more.
 We will discuss the implications of the latter in Section \@ref(limitations).
 We can also see that the number of NA responses rises from about 530 in the case of the income at age 16 variable to approximately 900 in the income variable.
-This seems counter intuitive as it would appear that less people are uncertain about their family income at age 16 than they are.
+This seems counter intuitive as it would appear that less people are uncertain about their family income at age 16 than they are now.
 We suspect however that this is not true.
@@
 On the other hand, someones income at age 16 is psychologically less of importance to them and carries no burden in their life, as it is not an expected metric of success years down the line.
-We also recognize that among men and women, there is a general almost a 50:50 split at each level of income but we do notice that for income level NA, a lot of people chose not to identify their gender.
-It would be interesting in another analysis to analyse possible evidences of gender pay gap and how people who do not fit in the *traditional* gender roles are affected by the advent of the American Dream and whether or not their position in life hinders their ability to increse in wealth over time.
+We also recognize that among men and women, there is in general almost a 50:50 split at each level of income but we do notice that for income level NA, a lot of people chose not to identify their gender.
+It would be interesting in another analysis to analyse possible evidences of gender pay gap and how people who do not fit in the *traditional* gender roles are affected by the advent of the American Dream and whether or not their position in life hinders their ability to increase in wealth over time.
 
-As we speak about an increase in wealth over time, we next want to see if the birth place of ones grandparents has any impact on income due to the so called *generational wealth*.
+As we speak about an increase in wealth over time, we next want to see if the birth place of ones grandparents has any impact on income due to so called *generational wealth*.
 

### 2022-03-21 outputs/paper/paper.Rmd
9th commit

@@
 # Data
-
 In this Data Section \@ref(data), we will provide a look into the data acquisition and processing methodology as well as a deep dive into the contents of the data.
 First of, (Table \@ref(tab:tableofdata)) give us a glimpse of the data.
+```{r echo = F}
+#table of variables that i am concerned with
+data_table <- reduced_data[,-c(1,6,8)]
 
-```{r echo = F}
-#table of variables that I am concerned with
-
-reduced_data[complete.cases(reduced_data), ] %>%
+data_table[complete.cases(data_table), ] %>%
    slice(1:7) %>%
@@
     caption = "GSS Cleaned Data",
-    col.names = c("Row Number", "Age", "Gender", "Born", "Marital Status", "Education", 
-                  "Divorce Status", "Degree", "Current Income", "Health Care", "Birth Year", 
-                  "First Born", "Income at 16", "Grandparents", "Category for Income at 16", 
-                  "Caegory for Current Income"),
+    col.names = c("Age", "Gender", "Born", 
+                  "Education", 
+                  "Birth Year", "Grandparents", 
+                  "Income at 16","Current Income"),
     digits = 1,
@@
     linesep = "")
+
 ```
@@
 # Visualizing the Data and The Implications
-In this analysis, I am interested at looking at how real the American Dream is and if it is found to be possible, whether or not there is any difference in attainment amoung genders.
+In this analysis, we am interested at looking at how real the American Dream is and if it is found to be possible, whether or not there is any difference in attainment amoung genders.
 As such, in this Section \@ref(visualizing-the-data-and-the-implications), we will create various visuals which will enable us to extract as much information as we can and we will speak about the implications that they make.
@@
 The first and most striking part of this visual is that for the current income, the highest threshold is $25,000 or more.
-We see that the both of our graphs are heavily skewed to wards higher incomes.
-I suspect that this is mainly a production of the bin widths decided on for the analysis.
+We see that the both of our graphs are heavily skewed to wards higher incomes.wesuspect that this is mainly a production of the bin widths decided on for the analysis.
 
@@
 ## Implications 2
-In the above visuals, we can see that in all the income groupings for both age 16 and current income, the largest subgrouping within each is people who have no grandparents born in the United states.
+In the above visuals, we can see that in all the income groupings for both age 16 and current income, the largest sub grouping within each is people who have no grandparents born in the United states.
 The second largest group are usually those with 2 grandparents born in the United States.
-It is pretty logical that 2 grandparents follows after zero grandparents as it is easy to recognize that most people do not go abroad to marry and marry a friend or someone that they know locally, who is most likely to be born in the united states.
-We would speculate that a similar trend can be observed in other *immigrant* countries Like Canada or Sweeden.
-What is the most fascinating however is that the vast majority of people who answered the survey are first or second generation Americans, perhaps having grandparents or parents who immigrated to this country chasing the American Dream.
+It is pretty logical that 2 grandparents follows after zero grandparents as it is easy to recognize that most people do not go abroad to marry and marry a friend or someone that they know locally, who is most likely to be born in the United States.
+We would speculate that a similar trend can be observed in other *immigrant* countries like Canada or Sweden.
+What is the most fascinating however, is that the vast majority of people who answered the survey are first or second generation Americans, perhaps having grandparents or parents who immigrated to this country chasing the American Dream.
 
-In a similar fashion to what was explained in \@ref(implications-1), both of our graphs are heavily skewed towards higher income.
+In a similar fashion to what was explained in Section \@ref(implications-1), both of our graphs are heavily skewed towards higher income.
 In the case of current income, it is almost certainly because of the bin width decided on by the people who built the survey which we personally find many problems with.
@@
 # Limitations
-In this limitations section, I want to focus on the data which is provided to us and have a  look into why it might not be easy to take the implications mentioned above at face value
+In this limitations section,wewant to focus on the data which is provided to us and have a  look into why it might not be easy to take the implications mentioned above at face value.
 
@@
 The first major flaw I've found in the data is that the bins that both the income at age 16 and current income are recorded in are not all the same size.
-For example, in the case of the the first bin has a size of 16 years, the second bin from ages 16 to 24 has width 9, the next two have width 19 with the last ranging in theory from 65 to infinity.
-Although it is possible to understand why the maintainers of the data choose to have a demarcation point of the bins at 65 and 16, it is very hard to understand why they opted to divide the interior portion into varying bin sizes.
+For example, in the case of the current income variable, the the first bin has a size of \$1,000, the second bin has a width of \$1,999, the third returns to about \$1,000 and as we go higher up the echelon, we have seemingly random alterations in bin widths from \$1,000 to \$2,000 to \$5,000.
+This can cause serious problems in statistics and especially for purposes of analysis.
 
-We know in data science that when plotting bar charts, the bin width can have a big implication on the presentation of the data and the things that we can infer about its distribution.
-If the bin width is too small or too large, we might reduce or add characteristics to the data that aren't actually representative of it.
+We know in data science that when plotting bar charts, the bin width can have a big implications on the presentation of the data and the things that we can infer about its distribution.
+If the bin width is too small, too large or uneven, we might reduce or add characteristics to the data that aren't actually representative of it.
+This will cause us to over-fit or under-fit a model when the time comes to create them.
+The other problem with the bin widths is that they are different from the income at age 16 to the current income.
+This is the subject of the following Section 5.2.
 
 ## Why cut off income at $25,000?
-use the us labor statistics data
-## Third discussion point
+According to a news release published by the United States Bureau of Labor Statistics (see the document called wkyeng.pdf in the literature folder in inputs), the median weekly income of US citizens in the 4th quarter of 2021 (furthest from the start of the pandemic) is $1,010.
+This information is from that of 116.3 million full-time wage and salary workers.
+If we do some basic mathematics to approximate the annual earning, we get a figure of over \$52,000 dollars per year.
+As such, we are unable to understand why the bins for current income have been designed as such.
+This is a major flaw in the technique used by the surveyers to collect the data.
+In section \@ref(next-steps)
 

### 2022-03-21 scripts/Data Acquisition and Processing.R

@@ -33,17 +33,14 @@ reduced_data <-
   select(age, 
          sexnow1,
          born,
-         marital,
          educ,
-         divorce,
-         degree,
          income,
-         nathealy,
          cohort,
-         agekdbrn,
          income16,
          granborn) %>% 
   rename(sex = sexnow1)
+
+
 
 reduced_data$income16_cat <- reduced_data$income16
 reduced_data$income_cat <- reduced_data$income

### 2022-03-22 outputs/paper/paper.Rmd
tenth assist

@@
 First of, (Table \@ref(tab:tableofdata)) give us a glimpse of the data.
-```{r echo = F}
+```{r tableofdata, echo = F}
 #table of variables that i am concerned with
@@
     linesep = "")
-
 ```
@@
 
-## Key features & Pros And Cons of the Methodology and Survey
+## Key features Pros And Cons of the Methodology and Survey
 
@@
 This has implications on the finding of our analysis in Section \@ref(limitations) as we will have to consider if we are able to generalize our finding from the relatively small sample that we have to the population of all US citizens over 18 years old.
-The small sample size and extremely long survey are both weaknesses of the survey which we will discuss in more depth in Section \@ref(limitations).
+The small sample size and extremely long survey are both weaknesses of the survey which we will discuss in more depth in Section \@ref(small-sample-size).
 
@@
 ```
-In figure \@ref(fig:figure-2), we see the overlapping of the two graphs, the first of which is the same as figure \@ref(fig:figure-1) and the bottom one is the same in essence but with the information for current family income instead of at age 16.
+In figure \@ref(fig:figure2), we see the overlapping of the two graphs, the first of which is the same as figure \@ref(fig:figure-1) and the bottom one is the same in essence but with the information for current family income instead of at age 16.
 The first and most striking part of this visual is that for the current income, the highest threshold is $25,000 or more.
-We see that the both of our graphs are heavily skewed to wards higher incomes.wesuspect that this is mainly a production of the bin widths decided on for the analysis.
+We see that the both of our graphs are heavily skewed to wards higher incomes.
+We suspect that this is mainly a production of the bin widths decided on for the analysis.
 
@@
 ## Visuals 3
-```{r echo=FALSE, fig.cap="write a figure caption here", fig.height=10, fig.width=12, message=FALSE, warning=FALSE}
+```{r, echo=FALSE, fig.cap="write a figure caption here", fig.height=10, fig.width=12, message=FALSE, warning=FALSE}
 #family income at 16 grouped by where individual was born
@@
 # Limitations
-In this limitations section,wewant to focus on the data which is provided to us and have a  look into why it might not be easy to take the implications mentioned above at face value.
+In this limitations section,we want to focus on the data which is provided to us and have a  look into why it might not be easy to take the implications mentioned above at face value.
 
 ## Uneven Bin width for Income Groups.
-The first major flaw I've found in the data is that the bins that both the income at age 16 and current income are recorded in are not all the same size.
-For example, in the case of the current income variable, the the first bin has a size of \$1,000, the second bin has a width of \$1,999, the third returns to about \$1,000 and as we go higher up the echelon, we have seemingly random alterations in bin widths from \$1,000 to \$2,000 to \$5,000.
+The first major flaw I've found in the data is that the bins that both the of the current income variable, the the first bin has a size of \$1,000, the second bin has a width of \$1,999, the third returns to about \$1,000 and as we go higher up the echelon, we have seemingly random alterations in bin widths from \$1,000 to \$2,000 to \$5,000.
 This can cause serious problems in statistics and especially for purposes of analysis.
@@
 As such, we are unable to understand why the bins for current income have been designed as such.
-This is a major flaw in the technique used by the surveyers to collect the data.
-In section \@ref(next-steps)
+This is a major flaw in the technique used by the surveyors to collect the data.
+In section \@ref(next-steps), we will discuss ways that the survey could improve on this.
+
+## Small sample size
+As we have reported in Section \@ref(key-features-pros-and-cons-of-the-methodology-and-survey), out of the 27,591 lines of sample (people who had been prompted to answer the questionnaire), 4,032 responses were collected.
+This means that for the 2021 GSS, the response rate was at 17.4 percent (see calculation on page 28 of the "GSS 2021 Codebook R1b.pdf").
+The United States has a population of over 330 million and even if all the people who had been prompted to answer the survey had responded it, the survey would only represent less than $8.36 \times 10^{-3}$ percent of the population.
+It is quite reasonable that as such, it would be hard to make generalizations for the entire adult united states population.
 
 ## Next steps
-
-Weaknesses and next steps should also be included.
+Following on from the latter section, we can make some pretty good suggestions on ways to improve the survey and subsequently, the data and the implications of the findings, so that they will be less ambiguous and more ethical to use with confidence.
+First off, having the data in predetermined bins is problematic from the perspective of the data analyst.
+We can start by creating even bins that respondents can answer but we can also allow the respondents to provide their exact income.
+The latter solution will give us more data, and it gives more options to the data analysis to conduct a reproducible analysis.
+In this manner, the data scientist who gets a hold of the data will have the liberty to conduct his analysis and chose the bins which work best for the task he is working on, instead of being obstructed by barriers built into the data.
 

### 2022-03-22 outputs/paper/paper.Rmd
11th
LAST COMMIT

@@
 # install.packages("tinytex")
-# tinytex::install_tinytex()
+tinytex::install_tinytex()
+#install.packages("kabeExtra")
 library(tinytex)
@@
 library(lubridate)
-
+library(kableExtra)
 reduced_data <- read_csv(here::here("inputs/data/reduced_data.csv"))
@@
 
+\newpage
 # Introduction
 
-You can and should cross-reference sections and sub-sections.
-For instance, Section \@ref(data).
-R Markdown automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point).
+If you ask a group of first graders what they aspire to be when they grow up, it wouldn't be surprising to hear a lot of kids say that they would want to be the richest person in the world.
+For as long as humanity has existed, the allure and pride associated with wealth and the accumulation of large sums of wealth has always been at the forefront of the social and economic paradigm.
+Whether it was Mansa Musa in East Africa in the 14th century or Jeff Bezos today, the wealthy have always stood out in large crowds.
+This aspect of the social sphere is one that can undoubtedly be found in American livelihood.
+It would be hard to deny for any of us that money rules the world, regardless of how we feel about that.
+We need it to put food on our tables, clothes on our backs and seemingly to increase our quality of life.
+As such, it would appear as if a surplus of money in ones possession could do no harm.
 
+For the last 50 years, people from all over the world have began to flock to the United States as immigrants in search of the so called American Dream.
+The American Dream is a Utopian idea in which the United States of America is a safe haven for all those looking to improve their quality of life via monetary gain.
+It implies that anyone (yes, absolutely anyone) who is willing to put in the hard work required to be successful, will have much improves abilities to make something of themselves in the United States than they could elsewhere, all else being equal.
+
+In this analysis, we got our hands on some of the 2021 United States General Social Survey data which is described with great extent in the Data Section \@ref(data).
+In our analysis, we decided to focus on how the various factors and trends in the data in order to try and gather some implications on the veracity of the american dream.
+To further break down the analysis, we decided to look into the distribution of the aforementioned data by segregating the data not only by the current income of respondents but also by their income at age 16, and trying to see if we can spot any relation between an increase in income from age 16 to now and the gender, or the number of grandparents of the respondent, born in the United States.
+
+In this paper, we will attempt to take you on a journey to explore the data that was used to produce the findings, we will visually present to you the findings and we will speak about the implications of the findings and later on in the Limitations Section \@ref(limitations), we will discuss the limitations of our data and problems that this may have when wanting to extract true, ethical, unbiased information from the implications of the findings.
+
+\pagebreak
 # Data
@@
 First of, (Table \@ref(tab:tableofdata)) give us a glimpse of the data.
-```{r tableofdata, echo = F}
+```{r tableofdata, echo = F, fig.pos = "!H"}
 #table of variables that i am concerned with
@@
     booktabs = TRUE,  
-    linesep = "")
+    linesep = "") %>% 
+  kable_styling(latex_options = "HOLD_position")
+
 ```
@@
 
+\pagebreak
 # Visualizing the Data and The Implications
@@
 ## Visual 1
-```{r figure 1, fig.cap="write a figure caption here", echo = FALSE, fig.width=12, fig.height=10, message = F, warning = F}
+```{r figure-1, fig.cap="write a figure caption here", echo = FALSE, fig.width=11, fig.height=9, message = F, warning = F}
 income16_gender <- ggplot(mapping = aes(x = factor(income16_cat, level = level_income16), group = sex, fill = sex), data = reduced_data) +
@@
 ```
-In this graph (Figure \@ref(fig:figure-1)), we can see the proportions of family income when the responded was 16 years of age.
+In this graph (Figure 1), we can see the proportions of family income when the responded was 16 years of age.
 We also see that an over-whelming majority of people in our sample had an income of over $40,000.
@@
 We would like to contrast this with the current income of responds. Below is the same graph as above, but we've also included a similar graph  for family income at the time of response.
-```{r figure2, fig.cap="write a figure caption here", echo = FALSE, fig.width=12, fig.height=10, message = F, warning = F}
+```{r figure2, fig.cap="Comparing proportions of current income and income at age 16, grouping by gender", echo = FALSE, fig.width=12, fig.height=10, message = F, warning = F}
 #income16 vs income grouped by gender
@@
 ```
-In figure \@ref(fig:figure2), we see the overlapping of the two graphs, the first of which is the same as figure \@ref(fig:figure-1) and the bottom one is the same in essence but with the information for current family income instead of at age 16.
+In Figure 2, we see the overlapping of the two graphs, the first of which is the same as Figure 1 and the bottom one is the same in essence but with the information for current family income instead of at age 16.
 The first and most striking part of this visual is that for the current income, the highest threshold is $25,000 or more.
@@
 As such, we will attempt to visualize the impact of grandparent immigration to the United States on the income of Americans.
-```{r, fig.cap="write a figure caption here", echo = FALSE, fig.width=12, fig.height=10, message = F, warning = F}
+```{r, fig.cap="Comparing proportions of current income and income at age 16, grouping by number of US born grandparents", echo = FALSE, fig.width=12, fig.height=10, message = F, warning = F}
 #fincome16 vs income grouped by where grandparents were born
@@
 ## Visuals 3
-```{r, echo=FALSE, fig.cap="write a figure caption here", fig.height=10, fig.width=12, message=FALSE, warning=FALSE}
+```{r, echo=FALSE, fig.cap="Comparing proportions of current income and income at age 16, grouping by US born grandparents or not", fig.height=10, fig.width=12, message=FALSE, warning=FALSE}
 #family income at 16 grouped by where individual was born
@@
 In this manner, the data scientist who gets a hold of the data will have the liberty to conduct his analysis and chose the bins which work best for the task he is working on, instead of being obstructed by barriers built into the data.
+To account for some of these changes, we have created a supplementary survey which collects a lot of the data we were interested in for this analysis.
+It stands independently from the 2021 US GSS.
+See the Appendix for the survey and additional information.
 

### 2022-03-22 outputs/paper/paper.Rmd
FINAL COMMIT

@@
 date: "`r format(Sys.time(), '%d %B %Y')`"
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "The American Dream is the idea that, as a nation, the United States embodies freedom that includes financial success, upward social mobility for families and children, as well as freedom to live as they please in a society with few barriers.
+We obtained and analyzed data from the 2021 United States General Social Survey, focusing on the relationship between the distribution of past and present income and various external demographic categories.
+We find that the data’s limitations limit our findings, where the analysis points to a heavy skew in income across all demographics however, the uneven bin width, income cutoff and small sample size obstruct us from taking our implications at face value.
+Although our findings have implications for the theory of the American Dream, due to the limitations we choose to create a stand alone supplementary survey that will aid us for our purposes."
 output:
@@
 This aspect of the social sphere is one that can undoubtedly be found in American livelihood.
-It would be hard to deny for any of us that money rules the world, regardless of how we feel about that.
+It would be hard fro any of ud to deny that money rules the world, regardless of how we feel about that.
 We need it to put food on our tables, clothes on our backs and seemingly to increase our quality of life.
@@
 For the last 50 years, people from all over the world have began to flock to the United States as immigrants in search of the so called American Dream.
-The American Dream is a Utopian idea in which the United States of America is a safe haven for all those looking to improve their quality of life via monetary gain.
+The American Dream is a Utopian idea in which the United States of America is a safe haven for all those looking to improve their quality of life, via monetary gain.
 It implies that anyone (yes, absolutely anyone) who is willing to put in the hard work required to be successful, will have much improves abilities to make something of themselves in the United States than they could elsewhere, all else being equal.
@@
 In this analysis, we got our hands on some of the 2021 United States General Social Survey data which is described with great extent in the Data Section \@ref(data).
-In our analysis, we decided to focus on how the various factors and trends in the data in order to try and gather some implications on the veracity of the american dream.
+In our analysis, we decided to focus on  the various factors and trends in the data in order to try and gather some implications on the veracity of the American Dream.
 To further break down the analysis, we decided to look into the distribution of the aforementioned data by segregating the data not only by the current income of respondents but also by their income at age 16, and trying to see if we can spot any relation between an increase in income from age 16 to now and the gender, or the number of grandparents of the respondent, born in the United States.
 
-In this paper, we will attempt to take you on a journey to explore the data that was used to produce the findings, we will visually present to you the findings and we will speak about the implications of the findings and later on in the Limitations Section \@ref(limitations), we will discuss the limitations of our data and problems that this may have when wanting to extract true, ethical, unbiased information from the implications of the findings.
+In this paper, we will attempt to take you on a journey to explore the data that was used to produce the findings, we will visually present to you the findings and we will speak about the implications of the findings and later on in the Limitations Section \@ref(limitations), we will discuss the limitations of our data and problems that this may have, when wanting to extract true, ethical, unbiased information from the implications of the findings.
 
@@
 When we tried to follow typical procedure to upload the 2021_stata.zip file into under data in the inputs folder, we received a message we had never received before.
-Figure 1 is a screen capture of that error.
-![The Error received upon attempting to upload the zip folder](~/paper_3_team_paper/inputs/2021_stata_error_message.png "The error message we received when trying to upload the zip folder.")
+Figure 1 bellow is a screen capture of that error.
+
+![The Error received upon attempting to upload the zip folder](~/paper_3_team_paper/inputs/2021_stata_error_message.png "The error message we received when trying to upload the zip folder."){.center}
 
@@
 # Appendix {-}
+Below is a link and a QR code for the supplementary survey alluded to in Section \@ref(next-steps).
+
+LINK TO SURVEY: https://docs.google.com/forms/d/1a3tK-uSkl0MDrjqxWW1aCKhLowQxdqBvE6Bid6wN3YY/edit?ts=6239fa2b
+
+![Supplimentary Survey](~/paper_3_team_paper/inputs/Supplimentary_survey_QR.png){width="300" height="500"}
 

### 2022-03-22 scripts/Data Acquisition and Processing.R

@@ -1,4 +1,4 @@
-#### Preamble dksbcasvkfvkcssf####
+#### Preamble ####
 # Purpose: Prepare the 2021 GSS data
 # Author: Adam Labas and Krupali Bhavsar
 # Data: March 20 2022

### 2022-03-22 outputs/paper/paper.Rmd
FINAL COMMIT

@@
 ---
-title: "Income Has Serious Implications On Many Facets of American Life in 2021"
+title: "Income Has Serious Implications On Many Facets of American Life in 2021."
 subtitle: "A Look Into The Implications of Americans' Incomes Reveals Information On The Veracity of American Dream."
